Notes on categorical logic and term calculi

### 2025-10-31 term-calculus/cat_logic_notes.tex
type theory update

@@

We call a multicategory representable if it is equipped with a chosed unit object and for every pair of objects with their tensor.

@@

>>>>

Pictorially these properties can be displayed as
\[
\tikzfig{./figures/composition_assoc}
\]

and therefore we have that $1 \leq i \leq k$ and $1 \leq j \leq k + m$.
Suppose, $j < i$ in which case we insert $f$ into $h$ avoiding $g$ and get $(h \circ_j f) \circ_{i + k - 1} g$.
If $i \leq j \leq i + m$, we have the same diagram just with different parentheses $h \circ_i (g \circ_{j - i + 1} f)$.
Similarly, we have that when $j \geq i + m$ then $(h \circ_{j - m + 1} f) \circ_{i} g$.
When we phrase the composition in terms of a substitution, these become the associativity and the interchange properties.
\begin{definition}

@@

We call a multicategory representable if it is equipped with a chosen unit object and for every pair of objects with their tensor.

@@

\begin{theorem}
There is an equivalence $\catname{RepMCat} \simeq \catname{MonCat}$.
\textcolor{red}{Can I define \textit{semilattice-enriched} $\catname{RepMCat}$ and restate the theorem above?}
\end{theorem}  
\begin{remark}
  We will eventually need to show that the type theory above defines a suitable multicategory and that this multicategory is representable (and therefore is equivalent to some monoidal category).
  The latter follows by the following argument. 
  We have a unit given by $\star$ and for each pair of objects we have their tensor product given by $\{-,-\}$.
  For the latter to be a tensor we need to make sure that we have a morphism $\chi : (A_1, \ldots, A_n) \to \otimes_i A_i$.
  It is given by the following derivation
  \[
  \inferrule*[right=$\chi$]{x_1 : A_1 \vdash x_1 : A_1 \ldots x_n : A_n \vdash x_n : A_n}{x_1 : A_1, \ldots, x_n : A_n \vdash \{x_1,\ldots,x_n\} : A_1 \otimes \ldots \otimes A_n}
  \]
  We need to show that
  \[
  \mathcal{M}(B_1,\ldots,B_k, \otimes_i A_i, C_1, \ldots, C_m; D) \xrightarrow{\sim} \mathcal{M}(B_1, \ldots, B_k,A_1, \ldots, A_n,C_1,\ldots, C_m ; D)
  \]
  Suppose we have a morphism from the left-hand side that is given by the derivation
  \[
  \inferrule*[right=$f$]{\ldots}{B_1, \ldots, B_k, z : \otimes_i A_i, C_1, \ldots, C_m} \vdash D
  \]
  and we do $f[\chi / z] : B_1, \ldots, B_k, A_1, \ldots, A_n, C_1, \ldots, C_m \vdash D$.
  In the other direction we do
  \[
  \inferrule*[]{\ldots}{B_1,\ldots,B_k, \otimes_i A_i, C_1, \ldots, C_m \vdash \text{match}(z, x_1\;\ldots\;x_n . D(x_1, \ldots, x_n))}
  \]
  These are inverses by $\beta/\eta$ laws.
\end{remark}
